---
title: "brief_examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{brief_examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Brief Examples

aRtpack can be used create specified dataframes that will map art when fed into `ggplot2` functions:

For example, `square_data()` creates a data frame that maps a square on to a ggplot:
```{r sq-ex}
library(ggplot2)
library(artpack)

# Use the function to create a data frame #
df_square <-
  square_data(x = 0,
              y = 0,
              size = 5,
              color = "purple",
              fill = "black"
              )

# Feed it into a ggplot #
df_square |>
  ggplot(aes(x = x, y = y)) +
  geom_polygon(
    fill = df_square$fill,
    color = df_square$color,
    linewidth = 2
  ) +
  coord_equal()


```

<br>

---

`rotator` will mathematically "rotate" existing data points in a data frame:
```{r rotator}
library(ggplot2)
library(artpack)

original_square <- data.frame(
  x = c(0, 3, 3, 0, 0),
  y = c(0, 0, 3, 3, 0)
)

rotated_square <- rotator(
  data = original_square,
  x = x,
  y = y,
  angle = 120,
  anchor = "center"
)

ggplot()+
  geom_path(data = original_square,
            aes(x,y),
            color = "red") +
  geom_polygon(data = rotated_square,
               aes(x,y),
               fill = "green") +
  coord_equal()
```
<br>

---

aRtpack functions are designed to be used in any part of your workflow. Experiment for some cool results:
```{r compilation}
library(ggplot2)
library(purrr)
library(dplyr)
library(tibble)
library(artpack)

# Create a base square #
square <- square_data(x = 0, y = 0, size = 1, group_var = TRUE) 

# Create square specs to be iterated on #
n_square <- 50
scaler <-seq(1,5, length = n_square)
fills <- art_pals("imagination", n = n_square)
angles <- seq(0,360, length = n_square)
group_n <- group_numbers(1:n_square)

# Add a random transformation for a little razzle dazzle ✨
theta <- seq(0,2*pi, length = 250)

list_opts <- list(scaler,
                  fills,
                  angles,
                  group_n)

df <- pmap(list_opts, ~rotator(square |>
                                 mutate(x = (x + ..1),
                                        y = (y + ..1) ,
                                        fill = ..2,
                                        group = paste0(group,..4)
                                 ),
                               x = x, y = y, angle = ..3)
) |> 
  list_rbind() |>
  mutate(x = x * cos(theta) + x,
         y = y * sin(theta))



df |>
  ggplot(aes(x = x, y = y, group = group)) +
  theme_void()+
  theme(plot.background = element_rect(fill = "#000000")) +
  geom_polygon(
    fill = df$fill,
    color = "#000000", 
    alpha = .5
  ) +
  coord_equal(expand = FALSE)

```

